---
title: "FINAL PROJECT STAT 575"
author: "Dominic Essuman"
date: "6/28/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}



# function to compute cauchy c(0,1) pdf at x using normal with mean 0 variance 2 as th e proposed dist
f <- function(x) {

 1/(pi*(1 + x^2))    
}

set.seed(1234)
m <- 10000 # chain length    
#sigma <- sqrt(2)     
x <- numeric(m)     
x[1] <- rnorm(1, mean = 0,sd =sqrt(2)) # generate first obs in the chain from candidate g() 
k <- 0     # counter for number of rejected values  
u <- runif(m)
for (i in 2:m) {
    # at each transition, the candidate point y
    # is generated from N(0,sqrt(2)
	xt <- x[i-1]
	y <- rnorm(1) 
	num <- f(y) * dnorm(xt)  #y is evaluated at f; xt is evaltd at g given y
	den <- f(xt) * dnorm(y) #xt is evaluated at f; y is evaltd at g given xt 
    if (u[i] <= num/den) 
        x[i] <- y    # if y is accepted, assign y to next value in the chain
	  else {     
        x[i] <- xt   # else assign old value of x to the next value  
        k <- k+1     # k record number of y rejected            
		}       
	}
print(100*k/m)
```
About 15.17\% of the candidate points are rejected, so the chain is efficient.

To check whether the generated values are independent (necessary for a random sample), we computed the autocorrelation for 100 adjacent values (or equivalently 10000/100 = 100 lags). It is easier to just check plot the autocorrelation values.  To plot the autocorrelation, we use the function `acf` in the`VGAM` package in R.

```{r}
library(VGAM) # install VGAM package
# computes autocorrelation for 100 adjacent values
acf(x, col = "red", lag.max = 100, ylab = "autocorrelation")  
```

 Note the highly correlated values in the first part of the chain. This tells us that the chain has not converged yet. We refer  these early values as  **burnin sample**. 

 In practice, we usually discard this first burnin sample values. From the autocorrelation plot above, we see that the autocorrelation started to hover around 0 after the 1000th (or 10th lag) value. With this, we discard the first 2000 values in the chain. 

 The last 9000 values can now be treated as the generated values from the Rayleigh distribution.

```{r, fig.width=7, fig.height=7}
par(mfrow = c(2, 2))  
a <- ppoints(100)     
QR <- tan((a-1)*pi)  #quantiles of Cauchy     
Qx <- quantile(x, a)
qqplot(QR, Qx, xlab ="Cauchy Quantiles", 
			ylab ="Sample Quantiles",
			main = "Original accepted values") 
abline(0,1, col=2)

b <- 2001      #discard the burnin sample     
y <- x[b: m] 
Qy <- quantile(y, a)
qqplot(QR, Qy, xlab="Cauchy Quantiles", 
			ylab ="Sample Quantiles",
			main = "Minus the burnin sample") 
abline(0, 1, col = 2)

hist(x, breaks = 30, prob = T, col = "gray", main = "Original accepted values")
curve(dcauchy(x, scale=1), add=TRUE, col="blue")

hist(y, breaks = 30, prob = T, col = "gray", main = "Minus the burnin sample")
curve(dcauchy(x, scale=1), add=TRUE, col="blue")
```
# using t distribution with degree of freedom 20

```{r}
f <- function(x) {

 1/(pi*(1 + x^2))    
}

set.seed(1234)
m <- 10000 # chain length    
#sigma <- sqrt(2)     
x <- numeric(m)     
x[1] <- rt(1,df =20) # generate first obs in the chain from candidate g() 
k <- 0     # counter for number of rejected values  
u <- runif(m)
for (i in 2:m) {
    # at each transition, the candidate point y
    # is generated from t distribution with degree of freedom 20
	xt <- x[i-1]
	y <- rt(1,df=20) 
	num <- f(y) * dt(xt,df = 20)  #y is evaluated at f; xt is evaltd at g given y
	den <- f(xt) * dt(y, df =20) #xt is evaluated at f; y is evaltd at g given xt 
    if (u[i] <= num/den) 
        x[i] <- y    # if y is accepted, assign y to next value in the chain
	  else {     
        x[i] <- xt   # else assign old value of x to the next value  
        k <- k+1     # k record number of y rejected            
		}       
	}
print(100*k/m)
```
```{r}
library(VGAM) # install VGAM package
# computes autocorrelation for 100 adjacent values
acf(x, col = "red", lag.max = 100, ylab = "autocorrelation")  
```
```{r, fig.width=7, fig.height=7}
par(mfrow = c(2, 2))  
a <- ppoints(100)     
QR <- tan((a-0.5)*pi)  #quantiles of Cauchy     
Qx <- quantile(x, a)
qqplot(QR, Qx, xlab ="Cauchy Quantiles", 
			ylab ="Sample Quantiles",
			main = "Original accepted values") 
abline(0,1, col=2)

b <- 4001      #discard the burnin sample     
y <- x[b: m] 
Qy <- quantile(y, a)
qqplot(QR, Qy, xlab="Cauchy Quantiles", 
			ylab ="Sample Quantiles",
			main = "Minus the burnin sample") 
abline(0, 1, col = 2)

hist(x, breaks = 30, prob = T, col = "gray", main = "Original accepted values")
curve(dcauchy(x, scale=1), add=TRUE, col="blue")

hist(y, breaks = 30, prob = T, col = "gray", main = "Minus the burnin sample")
curve(dcauchy(x, scale=1), add=TRUE, col="blue")
```